The host galaxy of the short GRB~111117A at $z = 2.211$\thanks{Based on observations collected at the European Southern Observatory, Paranal, Chile, Program ID: 088.A-0051 and 091.D-0904.}

### 2017-06-02 GRB111117A.tex
Address Zachs comments
Phew. Closes #6

We use the spectroscopic redshift, as well as new imaging data to constrain the
nature of the host galaxy as well as physical parameters of the GRB. The rest
frame X-ray derived hydrogen column density, for example, is high compared to a
complete sample of sGRBs and seems to follow the evolution with redshift as
traced by the hosts of long GRBs (lGRBs). This is consistent with the proportion
of the sGRB population originating in late-type galaxies live in similar
environments to that of lGRB hosts.

@@

From the age of the universe at explosion time, an initial progenitor separation
of $a_0 < 3.2~R_\odot$ is required in the case of a binary neutron star system. 
This puts constraints on the progenitor system evolution up to the time
of explosion.
}

@@

There is now mounting evidence that most short-duration gamma-ray bursts (GRBs)
come from the merger of neutron stars (NSs), either with another NS, or a black
hole, due to their apparent association with kilonovae \citep{Barnes2013a,
	Tanvir2013b, Yang2015, Jin2016, Rosswog2016}. The absence of associated
supernovae in deep searches \citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b}
supports this idea and distinguishes the physical origin of sGRBs from their
long-duration counterparts. %\citep[albeit see also][]{Fynbo2006b, Valle2006, Gal-Yam2006}.

The host galaxies of short GRBs are diverse. They are more massive and less actively 
star-forming on average than long GRB hosts \citep{Fong2013b}, while in some cases, no host 
galaxy can be identified \citep{Berger2010a, Tunnicliffe2014}. Together with their position 
within their hosts \citep{Fong2013a}, this suggests a progenitor system that can be very
long-lived, and is associated with stellar mass rather than star-formation rate
\citep{Berger2014}. The median redshift for sGRBs is $z\sim0.5$ \citep{Berger2014}, and
because most of these measurements come from the associated hosts, it is arguably biased
towards lower redshifts.

@@

The total lifetimes of NS binaries depends on their initial separations and
subsequent inspiral times, and impacts the timing and distribution of the
enrichment of the ISM and subsequent stars and planets with heavy $r$-process
elements \citep{VandeVoort2015, Wallner2015,  Ji2016}. Some limits can be
calculated based on models of star-formation histories of, and the spatial
distribution of sGRBs in, their host galaxies \citep[][]{Berger2014}. The most
distant cosmological bursts, however, offer direct, hard limits.

In this \emph{Letter} we present the spectrum of the host galaxy of the short GRB~111117A ($T_{90}=0.46$~s) and measure its redshift to be $z=2.211$. This value is significantly higher 
than the previously estimated redshift based on photometric studies. 
We present the GRB's rest frame 
properties based on this new distance compared to previous analyses
\citep{Margutti2012,Sakamoto2013} and revisit the host properties derived from
the new solution to the SED fit. %While no optical afterglow was detected, the
% excellent localization from a detection
%of the X-ray afterglow by the \emph{Chandra X-ray Observatory} also allows
%us to discuss the positioning and environmental properties of this remarkably
%distant short GRB.

Throughout this \emph{Letter} we use the $\Lambda$CDM cosmology provided by
\citet{Planck2015} in which the universe is flat with $H_0 = 67.7$\,km\,s$^{-1}$
and $\Omega_m = 0.307$ and report all magnitudes in the AB system.

@@

Spectroscopic observations were carried out using the cross-dispersed echelle
spectrograph, VLT/X-shooter \citep{Vernet2011}, at four seperate epochs. The
burst was followed up 38 hours after the Burst Alert Telescope (BAT) trigger
under ESO programme 088.A-0051 (PI: Fynbo) and again later using a different ESO
program 091.D-0904 (PI: Hjorth). X-shooter can cover the wavelength range from
3000~\AA~to 24\,800~\AA~(21\,000~\AA~when the $K$-band blocking filter is used)
across three spectroscopic arms. The bias-correction, flat-fielding, order
tracing, wavelength calibration, rectification, and flux calibration is then
carried out using the VLT/X-shooter pipeline version 2.8.4
\citep{Modigliani2010}. %Because the spectra are curved across each detector, a
% rectification algorithm
%is employed which introduces correlations between neighboring pixels. We select
%a pixel-scale of 0.2/0.2/0.6 \AA/pix for the UVB/VIS/NIR arm to minimize the
%degree of correlation while conserving the maximal resolution.
The observations are combined and extracted using scripts described in Selsing
et al. 2017 (in prep.) and available
online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}. The
signal-to-noise of the continuum in the near-infrared arm is too low to use the
optimal extraction algorithm \citep{Horne1986} and therefore the extraction is
carried out with a simple aperture. An overview of the spectroscopic
observations is given in Table~\ref{tab:spec_overview}, and the position of the
slit on the target is shown in Fig.~\ref{fig:spec_setup}.We show the extracted
spectrum in Fig.~\ref{fig:SED}.

We determine a redshift of $z = 2.211$ from the simultaneous detection of
emission lines interpreted as \lya, \oii, \hb, \oiii$\lambda$5007, and \ha, with
\hb~detected at low significance ($\sim 3 \sigma$). We show \oiii$\lambda$5007
in the insert in Fig.~\ref{fig:spec_setup}.  \ha~is only visible in the first
epoch, due to the $K$-band blocking filter used for the remainder observations.
The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and show
significant velocity structure along the slit. A drop in the continuum to the
blue of the \lya~line further supports the inferred redshift.

@@

Using the luminosity of \ha, we can infer the star-formation rate (SFR) of a 
galaxy \citep{Kennicutt1998}. At the redshift of the GRB host, \ha~is observed at 
21\,000~\AA~where the night sky is very bright. In addition, several bright sky-lines 
intersect the line, making an accurate estimate of the \ha-flux difficult. 
We obtain a limit on the SFR by integrating the part of \ha~free of contamination and 
correcting for the missing flux using the line shape. After converting the 
\citet{Kennicutt1998} relation to a \citet{Chabrier2003} initial mass function using
\citet{Madau2014}, we derive a limit of $SFR > 7 M_\odot$~yr$^{-1}$. 
From the SED-fit (Sect.~\ref{SED}), and the detection of \lya, the host is
constrained to contain very little or no dust, which is why we do not apply 
a dust-correction to the measured \ha~flux here.

@@

%Because the projected distance does not change much between redshift 1.3 and
% 2.1, all conclusions of \citet{Margutti2012} and \citet{Sakamoto2013} relating
% to host offset are unaffected.

@@

In addition to the spectroscopy presented above, we imaged the field of
GRB~111117A in multiple broad-band filters using the VLT equipped with FORS2
($gRIz$ filters) and HAWK-I ($JHK_{\mathrm{s}}$ filters), long after the burst
has faded. These new data are complemented by a re-analysis of some of the
imagery used in \citet{Margutti2012} and \citet{Sakamoto2013} that are available
to us here (GTC $gri$-band, TNG $R$-band, and Gemini $z$-band). A log of the
photometric observations and measured brightnesses is given in
Table~\ref{tab:phot_overview}.

All data were reduced, analyzed and fitted in a similar manner as described in detail in
\citet{Kruhler2011a} and, more recently, in \citet{Schulze2016}. Very briefly, we use our own
\texttt{python} and IRAF routines to perform a standard reduction which includes
bias/flat-field correction, de-fringing (if necessary), sky-subtraction, and
stacking of individual images. The photometry of the host was tied against magnitudes 
of field stars from the SDSS and 2MASS catalogs in the case of $grizJHK_{\mathrm{s}}$ filters. 
For the $R$ and $I$-band photometry we used the color transformations of
Lupton\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. 
We convert all magnitudes into the AB system if necessary, and correct for the Galactic 
foreground of $E_{B-V}=0.027~\mathrm{mag}$.

The multi-color spectral energy distribution (SED) is fit by \citet{Bruzual2003}
single stellar population models based on a \citet{Chabrier2003} initial mass
function in \emph{LePhare} \citep{Ilbert2006}, where the redshift is fixed to
the spectroscopic value of $z=2.211$. The best model is obtained with an
unreddened galaxy template, and returns physical parameters of absolute
magnitude ($M_B=-22.0\pm0.1$\,mag), stellar mass ($\log(M_{\star}/M_\odot) =
9.9\pm0.2$), stellar population age ($\tau = 0.5_{-0.3}^{+0.5}$ Gyr) and
star-formation rate ($SFR_{\mathrm{SED}}=11_{-4}^{+9}
M_\odot\,\mathrm{yr}^{-1}$). We show the SED fit in Fig.~\ref{fig:SED}.

Noteworthy is the discrepancy of our new VLT/FORS2 photometry and the
re-analysis of Gemini data to the $z$-band measurements of \citet{Margutti2012} and
\citet{Sakamoto2013}. Both of these authors report $z$-band photometry that is brighter 
by 0.8~mag to 1.0 mag than our value, whereas data taken in bluer filters are in excellent agreement. In fact, the large $i-z$ color is mistakenly interpreted as a 4000\,\AA\,break
driving the galaxy photometric redshift of the earlier works. Using
the revised photometry from Table~\ref{tab:phot_overview}, the photometric redshift of the
galaxy is $z_{\mathrm{phot}}=2.04_{-0.21}^{+0.19}$, consistent with the
spectroscopic value at the 1~$\sigma$ confidence level.

@@

We retrieved the automated data products provided by the \textit{Swift}-XRT GRB
repository\footnote{\url{http://www.swift.ac.uk/xrt\_products/00507901}}
\citep{Evans2009}. The X-ray afterglow light curve can be fit with a single power-law decay with
index $\alpha=1.27_{-0.10}^{+0.12}$. We performed a time-integrated spectral
analysis using data in photon counting (PC) mode in the widest time epoch where
the $0.3-1.5\,\mathrm{keV}$ to $1.5-10\,\mathrm{keV}$ hardness ratio is constant
(namely, from $t-T_0 = 205$~s to $t-T_0 = 203.5$~ks, for a total of 29.1~ks of
data) to prevent spectral changes that can affect the X-ray column density
determination. The obtained spectrum is well described by an absorbed power-law
model. The best-fit spectral parameters are a photon index of $2.1 \pm 0.4$ and
an intrinsic hydrogen column density $N_{\mathrm{H}}$ of $2.4_{-1.6}^{+2.4}
\times 10^{22}$~cm$^{-2}$ ($z=2.211$), assuming a Galactic $N_{\mathrm{H}}$ in
the burst direction of $4.1 \times 10^{20}$~cm$^{-2}$ \citep{Willingale2013}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Reinterpretation of restframe properties}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB~111117A securely 
belongs the short class of GRBs. Because the observed classification
indicators, $T_{90}$ and hardness ratio, do not depend strongly on redshift
\citep{Littlejohns2013a}, the updated redshift does not change this designation
significantly. 
%The intrinsic luminosity increases, as visible in the X-ray light curve
%in Fig. \ref{fig:sxray_lightcurve}, but it still is sub-luminous compared to
%the majority of long GRBs. 
\citet{Bromberg2013} investigated the degree to which the
long and short population distributions overlap and quantified the certainty in
the class membership. GRB~111117A has $96_{-5}^{+3}$ percent probability of
being short. Compared to the other two sGRB candidates at high redshift, GRB~060121
\citep{DeUgartePostigo2006, Levan2006} at $1.7 \lesssim z \lesssim 4.5$
($17_{-15}^{+14}$ per cent) and GRB~090426 \citep{Antonelli2009, Levesque2010,
	Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$ per cent), the certainty in class
membership for GRB~111117A is much higher.

Additionally, \citet{Horvath2010} classifies both GRB~060121 and GRB~090426 as
intermediate duration bursts. This comes from both events having very soft
spectra, as compared to the hard ones typically seen in sGRBs. Intermediate
bursts are very clearly related in their properties to long GRBs
\citep{DeUgartePostigo2011}, so they are unlikely to come from compact object
mergers.

@@

We plot the recalculated $N_\mathrm{H}$ in Fig.~\ref{fig:NH_z} where we compare with the
distributions of complete samples of both long and short GRBs. The lGRB sample
is from \citet{Arcodia2016} and the sGRB sample is from \citet{DAvanzo2014a}.
17 of the 99 long bursts do not have redshifts and likewise 5 out of 16 for
the short sample. Bursts without redshifts have been excluded for both groups.
GRB~111117A occupies a unique position in Fig.~\ref{fig:NH_z} with the highest
$N_H$ of all short burst. The short sample, excluding GRB~111117A, is located at
low redshifts ($z < 1$) and is found to populate a similar column density
environment to long GRBs at similar redshifts \citep{DAvanzo2014a}. The inferred
hydrogen column for GRB~111117A seems to follow the trend with increasing $N_H$
as a function of redshift as found for the long GRB afterglows \citep{Arcodia2016}.

@@

%From the clear host association, GRB~111117A does not belong to the hostless class of GRBs \citep{Berger2010a} and because the host exhibits emission lines indicative of a population of relatively young stars, 
As the majority of short GRBs \citep{Fong2013b}, the host of GRB~111117A is a
late-type galaxy and is entirely consistent in terms of stellar mass and stellar
age with the general population of short GRB hosts ($\left\langle M _*
\right\rangle = 10^{10.1} M_{\odot}$ and $\left\langle \tau _* \right\rangle =
0.3 $~Gyr) \citep{Leibler2010}. Being a late-type host, both the stellar mass
and sSFR are entirely within the range expected for the hosts of sGRBs
\citep{Behroozi2014}. The SFR is $\sim$1 order of magnitude higher than the
typical SFR for short GRB hosts galaxies \citep{Berger2014} and more similar to
the SFRs found in the hosts of long GRBs at a corresponding redshift
\citep{Kruhler2015}. The high SFR is partly a selection effect, because a less
star-forming galaxy would exhibit weaker emission lines, thus making the
redshift harder to determine. Additionally, it is natural to expect some
evolution in the hosts of sGRBs with redshift as illustrated in Sect.~\ref{restnH}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Implications for redshift distribution of short GRBs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A single sGRB at high redshift does little in terms of constraining the redshift
distribution and therethrough the progenitor channels. In particular, other
short-GRB hosts could be missed because they are intrinsically fainter and thus
this high-$z$ event is only detected due to the brightness of the host.
\citet{Berger2014} compiled a sample of sGRB host luminosities, normalized by
the characteristic galaxy luminosity at their respective redshift,
$L_B/L^{\star}_{B}$. 26 out of 39 hosts (66 per cent) in the sample have
redshifts. To convert the SED-inferred $M_B$ of GRB~111117A to
$L_B/L^{\star}_{B}$, we use the characteristic absolute B-band magnitude of the
Schechter function for the blue galaxies ($U - V < 0.25$) in the redshift window
$2.0 \leq z \leq 2.5$ from \citet{Marchesini2007} and find $L_B/L^{\star}_{B} =
1.2$, which is brighter than 70 per cent of the hosts in \citet{Berger2014} with
measured $L_B/L^{\star}_{B}$.

@@

A fraction of the bursts missing redshift are host-less and is therefore likely
at moderate redshifts \citep{Tunnicliffe2014}, but should some of the remainder
be at high redshift, the missed fraction will increase. If we assume that
\textit{all} the bursts that are missing redshifts in \citet{Berger2014} are at
high-$z$ and missed due to host faintness, 22 out of 39 hosts (55 per cent)
would be missed at $z = 2.211$. This serves as an upper limit on the fraction of
missed burst at high-$z$ and illustrates that we are likely \textit{not} missing a
large fraction of sGRBs redshift at $z \approx 2$, due to host faintness.

The theoretical redshift distribution of sGRBs depends on the type of delay-time
function used to model the progenitor system. The redshift of GRB~111117A puts
constraints on the type of delay-time models suitable for modeling. The
likelihood preferred lognormal time delay models investigated in
\citet{Wanderman2015} predicts a rate of sGRBs at $z = 2.211$, $\sim$ two orders
of magnitude lower compared to peak rate ($z = 0.9$). It is stated in
\citet{Wanderman2015} that this preference depends critically on the absence of
non-collapsar sGRB at $z \gtrsim 1.2$. The redshift of the burst, on the other
hand, is close to the expected peak in sGRB rate calculated for the power law
models \citep{Behroozi2014, Wanderman2015}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Constraints on the progenitor separation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

We use the spectroscopic redshift, as well as new imaging data to constrain the
nature of the host galaxy as well as physical parameters of the GRB. The rest-frame X-ray derived hydrogen column density, for example, is high compared to a
complete sample of sGRBs and seems to follow the evolution with redshift as
traced by the hosts of long GRBs (lGRBs). This is consistent with the proportion
of the sGRB population originating in late-type galaxies, live in similar
environments to that of lGRB hosts.

@@

From the age of the universe at the time of GRB explosion, an initial neutron
star (NS) separation of $a_0 < 3.2~R_\odot$ is required in the case where the
progenitor is a pair of inspiralling NSs. This constraint excludes some of the
longest formation channels for this burst.

}

@@

There is now mounting evidence that most short-duration $\gamma$-ray bursts
come from the merger of NSs, either with another NS, or a black
hole, due to their apparent association with kilonovae \citep{Barnes2013a,
	Tanvir2013b, Yang2015, Jin2016, Rosswog2016}. The absence of associated
supernovae in deep searches \citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b}
supports this idea and distinguishes the physical origin of sGRBs from their
long-duration counterparts. %\citep[albeit see also][]{Fynbo2006b, Valle2006, Gal-Yam2006}.

The host galaxies of sGRBs are diverse. They are more massive and less actively 
star-forming on average than lGRB hosts \citep{Fong2013b}, while in some cases, no host 
galaxy can be identified \citep{Berger2010a, Tunnicliffe2014}. Together with their position 
within their hosts \citep{Fong2013a}, this suggests a progenitor system that can be very
long-lived, and is associated with stellar mass rather than star-formation rate (SFR)
\citep{Berger2014}. The median redshift for sGRBs is $z\sim0.5$ \citep{Berger2014}, and
because most of these measurements come from the associated hosts, it is arguably biased
towards lower redshifts.

@@

The total lifetimes of NS binaries depends on their initial separations and
subsequent inspiral times. The delay time from formation to explosion impacts
the timing and distribution of the enrichment of the ISM with heavy $r$-process
elements \citep{VandeVoort2015, Wallner2015,  Ji2016}. Some limits can be
calculated based on models of star-formation histories of, and the spatial
distribution of sGRBs in, their host galaxies \citep[][]{Berger2014}. The most
distant cosmological bursts, however, offer direct, hard limits.

In this \emph{Letter} we present a spectrum of the host galaxy of the short GRB~111117A ($T_{90}=0.46$~s) and measure its redshift to be $z=2.211$. This value is significantly higher 
than the previously estimated redshift based on photometric studies. 
We present the GRB's rest frame 
properties based on this new distance compared to previous analyses
\citep{Margutti2012,Sakamoto2013} and revisit the host properties derived from
the new solution to the SED fit. %While no optical afterglow was detected, the
% excellent localization from a detection
%of the X-ray afterglow by the \emph{Chandra X-ray Observatory} also allows
%us to discuss the positioning and environmental properties of this remarkably
%distant short GRB.
Throughout this \emph{Letter} we use the $\Lambda$CDM cosmology parameters provided by
\citet{Planck2015} in which the universe is flat with $H_0 = 67.7$\,km\,s$^{-1}$
and $\Omega_m = 0.307$. All magnitudes are given in the AB system.

@@

Spectroscopic observations were carried out using the cross-dispersed echelle
spectrograph, VLT/X-shooter \citep{Vernet2011}, at four seperate epochs. The
burst was observed 38 hours after the Burst Alert Telescope (BAT) trigger under
ESO programme 088.A-0051 (PI: Fynbo) and again later under ESO programme
091.D-0904 (PI: Hjorth). X-shooter covers the wavelength range from 3000~\AA~to
24\,800~\AA~(21\,000~\AA~when the $K$-band blocking filter is used) across three
spectroscopic arms. We carried out the bias-correction, flat-fielding, order
tracing, wavelength calibration, rectification, and flux calibration using the
VLT/X-shooter pipeline version 2.8.4 \citep{Modigliani2010}.
%Because the spectra are curved across each detector, a
% rectification algorithm
%is employed which introduces correlations between neighboring pixels. We select
%a pixel-scale of 0.2/0.2/0.6 \AA/pix for the UVB/VIS/NIR arm to minimize the
%degree of correlation while conserving the maximal resolution.
The observations are combined and extracted using scripts described in Selsing
et al. 2017 (in prep.) and available
online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}. 
%The
%signal-to-noise of the continuum in the near-infrared arm is too low to use the
%optimal extraction algorithm \citep{Horne1986} and therefore the extraction is
%carried out with a simple aperture.
An overview of the spectroscopic observations is given in
Table~\ref{tab:spec_overview}, and the slit position is shown in
Fig.~\ref{fig:spec_setup}. We show the extracted spectrum in Fig.~\ref{fig:SED}.

We determine a redshift of $z = 2.211$ from the simultaneous detection of
emission lines interpreted as \lya, \oii, \hb, \oiii$\lambda$5007, and \ha. 
\hb~is detected at low significance ($\sim 3 \sigma$). We show \oiii$\lambda$5007
in the insert in Fig.~\ref{fig:spec_setup}.  \ha~is only visible in the first
epoch due to the $K$-band blocking filter used for the remainder observations.
The nebular lines exhibit a spatial extent of $\sim$ 1\farc5 and show
significant velocity structure along the slit. A drop in the continuum to the
blue of the \lya~line further supports the inferred redshift.

@@

Using the luminosity of \ha, we can infer the star-formation rate (SFR) of a 
galaxy \citep{Kennicutt1998}. At the redshift of the GRB host, \ha~is observed at 
21\,000~\AA~where the night sky is very bright. In addition, several bright sky-lines 
are superposed on the line, making an accurate estimate of the \ha-flux difficult. 
We obtain a limit on the SFR by integrating the part of \ha~free of contamination and 
correcting for the missing flux using the line shape. After converting the 
\citet{Kennicutt1998} relation to a \citet{Chabrier2003} initial mass function using
\citet{Madau2014}, we derive a limit of $SFR > 7 M_\odot$~yr$^{-1}$. 
From the SED-fit (Sect.~\ref{SED}), and the detection of \lya, the host is
constrained to contain very little or no dust, which is why we do not apply 
a dust-correction to the measured \ha~flux here.

@@

@@

In addition to the spectroscopy presented above, we imaged the field of
GRB~111117A in multiple broad-band filters using the VLT equipped with FORS2
($gRIz$ filters) and HAWK-I ($JHK_{\mathrm{s}}$ filters), long after the burst
faded. These new data are complemented by a re-analysis of some of the imaging
used in \citet{Margutti2012} and \citet{Sakamoto2013} that are available to us
(GTC $gri$-band, TNG $R$-band, and Gemini $z$-band). A log of the photometric
observations and measured brightnesses is given in
Table~\ref{tab:phot_overview}.

All data were reduced, analyzed and fitted in a similar manner as described in detail in
\citet{Kruhler2011a} and, more recently, in \citet{Schulze2016}. Briefly, we use our own
\texttt{Python} and IRAF routines to perform a standard reduction which includes
bias/flat-field correction, de-fringing (if necessary), sky-subtraction, and
stacking of individual images. The photometry of the host was tied against magnitudes 
of field stars from the SDSS and 2MASS catalogs in the case of $grizJHK_{\mathrm{s}}$ filters. 
%For the $R$ and $I$-band photometry, we used the color transformations of
%Lupton\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. 
We convert all magnitudes into the AB system, and correct for a Galactic 
foreground of $E_{B-V}=0.027~\mathrm{mag}$.

The multi-color spectral energy distribution (SED) is fit by \citet{Bruzual2003}
single stellar population models based on a \citet{Chabrier2003} initial mass
function in \emph{LePhare} \citep{Ilbert2006}, where the redshift is fixed to
the spectroscopic value of $z=2.211$. The best fit model is an
unreddened galaxy template, and returns physical parameters of absolute
magnitude ($M_B=-22.0\pm0.1$\,mag), stellar mass ($\log(M_{\star}/M_\odot) =
9.9\pm0.2$), stellar population age ($\tau = 0.5_{-0.3}^{+0.5}$ Gyr) and
star-formation rate ($SFR_{\mathrm{SED}}=11_{-4}^{+9}
M_\odot\,\mathrm{yr}^{-1}$). We show the SED fit in Fig.~\ref{fig:SED}.

Noteworthy is the discrepancy of our new VLT/FORS2 photometry and the
re-analysis of the Gemini data to the $z$-band measurements of \citet{Margutti2012} and
\citet{Sakamoto2013}. Both of these authors report $z$-band photometry that is brighter 
by 0.8~mag to 1.0 mag compared to our value, where data taken in bluer filters are in excellent agreement. The large $i-z$ color is mistakenly interpreted as a 4000\,\AA\,break
driving the galaxy photometric redshift of the earlier works. Using
the revised photometry from Table~\ref{tab:phot_overview}, the photometric redshift of the
galaxy is $z_{\mathrm{phot}}=2.04_{-0.21}^{+0.19}$, consistent with the
spectroscopic value at the 1~$\sigma$ confidence level.

@@

We retrieved the automated data products provided by the \textit{Swift}-XRT GRB
repository\footnote{\url{http://www.swift.ac.uk/xrt\_products/00507901}}
\citep{Evans2009}. 
%The X-ray afterglow light curve can be fit with a single power-law decay with an
%index $\alpha=1.27_{-0.10}^{+0.12}$. We performed a time-integrated spectral
%analysis using data obtained in photon counting (PC) mode in the widest time epoch where
%the $0.3-1.5\,\mathrm{keV}$ to $1.5-10\,\mathrm{keV}$ hardness ratio is constant
%(namely, from $t-T_0 = 205$~s to $t-T_0 = 203.5$~ks, for a total of 29.1~ks of
%data) to prevent spectral changes that can affect the X-ray column density
%determination. 
The obtained spectrum is well described by an absorbed power-law
model and the best-fit spectral parameters are a photon index of $2.1 \pm 0.4$ and
an intrinsic hydrogen column density $N_{\mathrm{H}}$ of $2.4_{-1.6}^{+2.4}
\times 10^{22}$~cm$^{-2}$ ($z=2.211$), assuming a Galactic $N_{\mathrm{H}}$ in
the burst direction of $4.1 \times 10^{20}$~cm$^{-2}$ \citep{Willingale2013}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Reinterpretation of the restframe properties}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Because the projected distance does not change much between $z = 1.3$ and $z =
2.211$, all conclusions of \citet{Margutti2012} and \citet{Sakamoto2013}
relating to host offset are unaffected.

@@

As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB~111117A is securely 
classified as belonging to the short class of GRBs. Because the observed classification
indicators, $T_{90}$ and hardness ratio, do not depend strongly on redshift
\citep{Littlejohns2013a}, the updated redshift does not change this designation
significantly. 
%The intrinsic luminosity increases, as visible in the X-ray light curve
%in Fig. \ref{fig:sxray_lightcurve}, but it still is sub-luminous compared to
%the majority of long GRBs. 
\citet{Bromberg2013} investigated the degree to which the
long and short population distributions overlap and quantified the certainty in
the class membership. GRB~111117A has $96_{-5}^{+3}$ percent probability of
being a sGRB. Compared to the other two sGRB candidates at high redshift, GRB~060121
\citep{DeUgartePostigo2006, Levan2006} at $1.7 \lesssim z \lesssim 4.5$
($17_{-15}^{+14}$ per cent) and GRB~090426 \citep{Antonelli2009, Levesque2010,
	Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$ per cent), the certainty in class
membership for GRB~111117A is much higher.

%Additionally, \citet{Horvath2010} classifies both GRB~060121 and GRB~090426 as
%intermediate duration bursts. This comes from both events having very soft
%spectra, as compared to the hard ones typically seen in sGRBs. Intermediate
%bursts are very clearly related in their properties to lGRBs
%\citep{DeUgartePostigo2011}, so they are unlikely to come from compact object
%mergers.

@@

We show the recalculated $N_\mathrm{H}$ in Fig.~\ref{fig:NH_z} where we compare
with the distributions of complete samples of both long and short GRBs. The lGRB
sample is from \citet{Arcodia2016} and the sGRB sample is from
\citet{DAvanzo2014a}. 17 out of the 99 long bursts do not have redshifts, and
likewise for 5 out of 16 in the short sample. Bursts without redshifts have been
excluded for both groups. GRB~111117A occupies a unique position in
Fig.~\ref{fig:NH_z} with the highest $N_H$ of all short bursts. Additionally the
upper limit at $z = 2.609$ belongs to GRB~090426 and does not belong in the
short sample. The short sample, excluding GRB~111117A, is located at low
redshifts ($z < 1$) and is found to populate a similar column density
environment to lGRBs at similar redshifts \citep{DAvanzo2014a}. The inferred
hydrogen column for GRB~111117A seems to follow the trend with increasing $N_H$
as a function of redshift as found for the lGRB afterglows \citep{Arcodia2016}.

@@

%From the clear host association, GRB~111117A does not belong to the hostless class of GRBs \citep{Berger2010a} and because the host exhibits emission lines indicative of a population of relatively young stars, 
As the majority of sGRBs \citep{Fong2013b}, the host of GRB~111117A is a
late-type galaxy and is entirely consistent in terms of stellar mass and stellar
age with the general population of sGRB hosts ($\left\langle M _*
\right\rangle = 10^{10.1} M_{\odot}$ and $\left\langle \tau _* \right\rangle =
0.3 $~Gyr) \citep{Leibler2010}. Being a late-type host, both the stellar mass
and sSFR are entirely within the range expected for the hosts of sGRBs
\citep{Behroozi2014}. The SFR is $\sim$1 order of magnitude higher than the
typical SFR for sGRB host galaxies \citep{Berger2014} and more similar to
the SFR found in the hosts of lGRBs at a corresponding redshift
\citep{Kruhler2015}. The high SFR is partly a selection effect, because a less
star forming galaxy would exhibit weaker emission lines, thus making the
redshift harder to determine. Additionally, it is natural to expect some
evolution in the hosts of sGRBs with redshift as illustrated in Sect.~\ref{restnH}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Implications for the redshift distribution of sGRBs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A single sGRB at high redshift does little in terms of constraining the redshift
distribution of sGRBs. In particular, other sGRB hosts could be missed because
they are intrinsically fainter and thus this high-$z$ event is only detected due
to the brightness of it's host. \citet{Berger2014} compiled a sample of sGRB
host luminosities, normalized by the characteristic galaxy luminosity at their
respective redshift, $L_B/L^{\star}_{B}$. Out of the 39 hosts in the sample, 26
(66 per cent) have redshifts. To convert the SED-inferred $M_B$ of GRB~111117A
to $L_B/L^{\star}_{B}$, we use the characteristic absolute $B$-band magnitude of
the Schechter function for blue galaxies ($U - V < 0.25$) in the redshift window
$2.0 \leq z \leq 2.5$ from \citet{Marchesini2007}. For GRB~111117A, we obtain
$L_B/L^{\star}_{B} = 1.2$, which is brighter than 70 per cent of the hosts in
\citet{Berger2014} with measured $L_B/L^{\star}_{B}$.

@@

A fraction of the bursts missing redshift are host-less and is therefore likely
at moderate redshifts \citep{Tunnicliffe2014}, but should some of the remainder
be at high redshift, the missed fraction will increase. If we assume that
\textit{all} the bursts that are missing redshifts in \citet{Berger2014} are at
high-$z$ and missed due to host faintness, 22 out of 39 hosts (55 per cent)
would be missed at $z = 2.211$. This serves as an upper limit on the fraction of
missed burst at high-$z$ and illustrates that we are likely \textit{not} missing a
large fraction of sGRBs redshift at $z \approx 2$ due to host faintness.

The theoretical redshift distribution of sGRBs depends on the type of delay-time
function used to model the progenitor system. The likelihood preferred lognormal
time delay models investigated by \citet{Wanderman2015} predict a sGRB rate at
$z = 2.211$, $\sim$ two orders of magnitude lower compared to the peak rate at
$z = 0.9$. According to \citet{Wanderman2015}, this preference depends
critically on the absence of non-collapsar sGRBs at $z \gtrsim 1.2$. The
redshift of GRB~111117A, on the other hand, is close to the expected peak in
sGRB rate calculated using the power law delay time models \citep{Behroozi2014,
	Wanderman2015}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Constraints on progenitor separation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

At $z = 2.211$, the age of the universe is 3 Gyr. If the progenitor systems of
sGRBs are the merger of two NSs, this sets a hard upper limit to the coalescence
timescale for such a system. In the absence of other mechanisms, the timescale
of the orbital decay of the system is set by the energy loss due to
gravitational waves, which in turn is set by the mass of the constituent compact
objects and the separation of the two \citep{Postnov2014}. If we assume that the
formation timescale of the first galaxies is short compared to the time since
the Big Bang \citep{Richard2011}, and if we assume a mass of 1.4~$M_\odot$ for
each of the NSs at the time of system formation, this places a hard upper limit
on the initial separation, of $a_0 < 3.2~R_\odot$.

If we use the stellar population age from our SED fit, then we obtain a
(softer) limit on the initial separation of $a_0 < 2.1~R_\odot$. However, this
does not account for the possibility there could be an underlying stellar
population of older stars from a previous star-formation epoch. The delay time
between formation and explosion is well accommodated by the models of
\citet{Belczynski2006}, although the longest delay times are excluded.
This is especially true given the late type nature of the host
\citep{OShaughnessy2008}.
